Supervised Topic Modeling in Clinical Text

### 2011-05-10 paper/main.tex
Shuffled things around, updated model.

\section{Introduction}
\label{sec:introduction}
Despite the growing emphasis on meaningful use of technology in medicine, many aspects of medical record-keeping remain a manual process.  Diagnostic coding for billing and insurance purposes is often handled by professional medical coders who must explore a patient's extensive clinical record before assigning the proper codes.  So while electronic health records (EHRs) should be adopted by most medical institutions within the next several years, largely due to the provisions of HITECH under the American Recovery and Reinvestment Act \citep{Blumenthal2009}, there has been little movement forward in automating medical coding.

@@

While there have been some attempts to automatically classify groups of patients as a potential preliminary step to ICD-9 code assignment \citep{Ruch2008, FreitasJunior2006, RibeiroNeto2001, Brown2006}, fully automatic assignment of ICD-9 codes to medical text became a more prevalent research topic only in the last few years. A subset of earlier work proposed various methods on small corpora, based on a few specific diseases \citep{Rao2003} but the most recent and promising work on the subject was inspired by the 2007 Medical NLP Challenge: “International Challenge: Classifying Clinical Free Text Using Natural Language Processing” (website). Most of the classification strategies included word matching and rule-based algorithms. \citep{Goldstein2007, Crammer2007, Farkas2008}.  The data set given to the participants consisted only of documents that were 1-2 lines each and all of the documents were radiology reports - clearly limiting the scope of potential ICD-9 codes which could be assigned. The only paper which has attempted to work with a document scope as large as ours was the 2008 Lita et al publication \citep{Lita2008}.  Lita proposed support vector machine and bayesian ridge regression methods to assign appropriate labels to the documents but did not utilize the ICD-9 hierarchy to leverage more comprehensive predictions.

@@

@@

\subsection{Data}
Our data set was gathered from the clinical data warehouse of NewYork - Presbyterian Hospital. The data consisted of free-text discharge summaries and their respective ICD-9 codes. A discharge summary is a clinical report prepared by a physician or other health professional at the conclusion of a hospital stay or series of treatments. The note outlines the patient’s chief complaint, diagnostic findings, therapy administered, patient’s response to the chosen therapy, the treatment plan and the recommendations upon discharge. The ICD-9 codes used to structure the discharge summary data are part of a controlled terminology which is the international standard diagnostic classification for epidemiological, health management, and clinical purposes (http://www.who.int/classifications/icd/en/). The codes are classified in a rooted-tree structure, with each edge representing an is-a relationship between parent and child, such that the parent diagnosis subsumes the child diagnosis.  In the hospital, ICD-9 codes are generated manually by trained medical coders, who review all the information in the discharge summary.  For the purposes of sLDA, ICD-9 codes will be used as labels for discharge summaries.

\subsection{Pre-Processing}
Patient discharge summaries and their associated ICD-9 diagnoses are stored in two different places in the NewYork - Presbyterian data warehouse, and so had to be linked together before being fed into the sLDA algorithm.  Each discharge note and set of diagnoses were assigned a patient unique identifier (PUID) and a visit unique identifier (VUID), allowing the two types of data to be linked.

Preparation of the diagnostic codes involved inference over the ICD-9 hierarchy.  The is-a relationships of the hierarchy allowed us to make two important assumptions.  First, if a diagnosis was observed to be present, all of its ancestors could be assumed to be present as well (e.g., if a patient had malignant hypertension, it could be assumed that they also had essential hypertension.  Second, if a diagnosis was observed to be absent, it could be assumed that all of its descendants were also absent (e.g. if a patient did not have essential hypertension, it could be assumed that they did not have malignant hypertension).  Unfortunately, ICD-9 code observations never include observations of disease absence.  ICD-9 codes are only documented when the condition is observed to be present.  Additionally, ICD-9 codes are known to have relatively low sensitivity; conditions that are present are often not documented in a set of ICD-9 codes (Surjan, 1999). Given these facts, we made the following assumptions regarding each visit: recorded diagnoses and their ancestors were labeled as true; diagnoses that were observed at some time for a patient but not at the current visit were labeled as unobserved; and diagnoses that had never been listed for a patient were labeled as false for all of that patient’s visits. This last assumption captures the belief that parts of the ICD-9 hierarchy that are never observed for a particular patient are almost certain to be false.  Additionally, for computational purposes, we decided not to include an ICD-9 code at all if neither it nor one of its descendants had been assigned to a patient in any of the records in our dataset.

\subsection{Supervised Topic Models}
Latent dirichlet allocation (LDA) is a generative probabilistic model of corpora that represents documents as a mixed membership bag-of-words. Also known as topic models, these models infer the latent structure, or topics, of documents in a corpus. Each document is represented as a collection of words, generated from a set of topic assignments (one for each word), where each topic assignment is drawn from a distribution over topics \citep{Blei2003}.

@@

\subsection{Generative Model}

%
\begin{figure}[h]
%tbp] %  figure placement: here, top, bottom, or page
 \centering \includegraphics[width=4in]{model_1} \includegraphics[width=3in]{model_2}
\caption{adapted sLDA model}

Given the number of topics, K, the global prior over topic proportions,
$\alpha'$, and the prior over topics, $\gamma$, the generative process
for documents and responses is as follows: 
\begin{enumerate}
\item For each topic:

\begin{enumerate}
\item Draw a distribution over words $\beta_{k}\sim Dir\left(u,\gamma\right)$
\end{enumerate}
\item For each ICD9 Code:

\begin{enumerate}
\item Draw regression coefficient $\eta_{i}\mid\mu,\sigma\sim\mathcal{N}\left(\mu,\sigma\right)$ 
\end{enumerate}
\item Draw a prior over topic proportions $m\mid\alpha'\sim Dir\left(u,\alpha'\right)$ 
\item For each document:

\begin{enumerate}
\item Draw topic proportions $\theta_{d}\mid\alpha\sim Dir\left(m,\alpha\right)$ 
\item For each word:

\begin{enumerate}
\item Draw topic assignment $z_{n,d}\mid\theta_{d}\sim Mult\left(\theta_{d}\right)$ 
\item Draw word $w_{n,d}\mid z_{n,d},\beta_{1:K}\sim Mult\left(\beta_{z_{n}}\right)$ 
\end{enumerate}
\item For each ICD-9 code from the root of the hierarchy and recursively
descending the tree:

\begin{enumerate}
\item Draw a response variable $y_{i}\mid\bar{z},\eta_{i},y_{parent}\sim\Phi(\eta_{i}^{T}\bar{z},y_{parent})$
where $\bar{z}=N^{-1}\sum_{n=1}^{N}z_{n}$ and $\Phi$ refers to a
conditional probit model.
\end{enumerate}
\end{enumerate}
\end{enumerate}
We will employ a data augmentation scheme with auxiliary variables
$a_{i}$ in the probit model where:

\begin{equation}
y_{i}\sim\begin{cases}
\begin{array}{c}
1,\\
0,\end{array} & \begin{array}{c}
a_{i}>0\; and\; y_{parent}=1\\
otherwise\end{array}\end{cases}\end{equation}

\begin{equation}
a_{i}\sim\mathcal{N}\left(\eta_{i}^{T}\bar{z},1\right)\end{equation}

Given an observation of a set of ICD-9 codes and a document, the posterior
distribution for the latent variables is given by Equation 4.

The denominator for this distribution is the marginal probability
of the data and cannot be solved in closed form. This is often the
case in evaluating posterior distributions of non-trivial probabilistic
models. We will appeal to one of the common methods for approximating
posterior distributions in the face of intractable normalization factors:
Markov chain Monte Carlo (MCMC) sampling. Since in this model it is
possible to sample from the conditional distributions for all variables
we will use the Gibbs sampling algorithm to obtain an approximation
to this posterior. In particular, we will derive a collapsed Gibbs
sampler for the supervised topic model.

\begin{equation}
=\prod_{i=1}^{I}\left[p\left(\eta_{i};\mu\right)\prod_{m=1}^{M}p\left(y_{m,i}\mid z_{m,1:N},\eta_{i},\xi\right)\right]\prod_{k=1}^{K}\frac{\Gamma\left(\sum_{v=1}^{V}\beta_{v}\right)}{\prod_{v=1}^{V}\Gamma\left(\beta_{v}\right)}\frac{\prod_{v=1}^{V}\Gamma\left(n_{\left(\bullet\right),v}^{k}+\beta_{v}\right)}{\Gamma\left(\sum_{v=1}^{V}n_{\left(\bullet\right),v}^{k}+\beta_{v}\right)}\prod_{m=1}^{M}\frac{\Gamma\left(\sum_{k=1}^{K}\alpha_{k}\right)}{\prod_{k=1}^{K}\Gamma\left(\alpha_{k}\right)}\frac{\prod_{k=1}^{K}\Gamma\left(n_{m,\left(\bullet\right)}^{k}+\alpha_{k}\right)}{\Gamma\left(\sum_{k=1}^{K}n_{m,\left(\bullet\right)}^{k}+\alpha_{k}\right)}\end{equation}

\end{small} \end{landscape}

To derive the Gibbs sampler we evaluate the individual conditional
probability distributions for all unobserved variables.

\subsubsection{$p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{Y},\mathbf{w},\mathbf{\eta},\alpha,\beta,\mu,\xi\right)$}

For the purposes of sampling, we will be able to derive a representation
of the joint distribution isolating a particular latent variable,
$z$, for a word instance, $n$, in a document instance, $m$. The
conditional probability with respect to this latent variable is proportional
to the joint distribution up to a constant.

\begin{equation}
p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{Y},\mathbf{w},\mathbf{\eta},\alpha,\beta,\mu,\xi\right)\propto p\left(z_{m,n},\mathbf{z_{-\left(m,n\right)}},\mathbf{Y},\mathbf{w},\mathbf{\eta},\alpha,\beta,\mu,\xi\right)\end{equation}

Due to the factorization of this model we can rewrite the joint distribution
as the following:

\begin{equation}
\propto\prod_{i=1}^{I}\left[p\left(\eta_{i};\mu\right)p\left(y_{m,i}\mid z_{m,1:N},\eta_{i}\xi\right)\right]p\left(z_{m,n},\mathbf{z_{-\left(m,n\right)}},\mathbf{w},\mathbf{\eta},\alpha,\beta,\mu\right)\end{equation}

\begin{equation}
\propto\prod_{i=1}^{I}\left[p\left(y_{m,i}\mid z_{m,1:N},\eta_{i},\xi\right)\right]\left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\cdot\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\cdot\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}\end{equation}

Here, $n_{m,v}^{k,-\left(m,n\right)}$ represents the count of word
$v$ in document $m$ assigned to topic $k$ omitting the $(m,n)^{th}$
word count. For exponential family distributions, the normalization
contant, $h(y_{m,i})$, does not depend on $z_{m,n}$.

\begin{equation}
\propto\exp\left\{ \sum_{i=1}^{I}\left(\eta_{i}^{T}\bar{z}\right)y_{m,i}-A\left(\eta_{i}^{T}\bar{z}\right)\right\} \left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}\label{eq:z_likelihood}\end{equation}

Given this expression, $p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{Y},\mathbf{w},\mathbf{\eta},\alpha,\beta,\mu\right)$
can be sampled through enumeration as seen in Equation 13. In the
case of probit regression, the expression for \ref{eq:z_likelihood}
evaluates to Equation 14. Equivalently we can parameterize the model
with an auxiliary variable $a_{i}$, resulting in Equation 15.

%\begin{equation}
%p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{Y},\mathbf{w},\mathbf{\eta},\alpha,\beta,\mu\right)=\frac{\exp\left\{ \sum_{i=1}^{I}\left(\eta_{i}^{T}\bar{z}\right)y_{m,i}-A\left(\eta_{i}^{T}\bar{z}\right)\right\} \left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}}{\sum_{k=1}^{K}\exp\left\{ \sum_{i=1}^{I}\left(\eta_{i}^{T}\bar{z}\right)y_{m,i}-A\left(\eta_{i}^{T}\bar{z}\right)\right\} \left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}}\end{equation}

%\begin{equation}
%p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{w},\mathbf{\eta},\alpha,\beta,\mu\right)\propto\prod_{i=1}^{I}\left[\Phi\left(\eta_{i}^{T}\bar{z}\right)^{y_{m,i}}\left(1-\Phi\left(\eta_{i}^{T}\bar{z}\right)\right)^{1-y_{m,i}}\right]\left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}\end{equation}

%\begin{equation}
%p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{w},\mathbf{\eta},\mathbf{a},\alpha,\beta,\mu\right)\propto\prod_{i=1}^{I}\left[\delta\left(sign\left(a_{m,i}\right)=y_{m,i}\right)\mathcal{N}\left(a_{m,i}\mid\eta_{i}^{T}\bar{z,}1\right)\right]\left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}\end{equation}

\begin{landscape}

>>>>

\section{Introduction}

\label{sec:introduction} Despite the growing emphasis on meaningful
use of technology in medicine, many aspects of medical record-keeping
remain a manual process. Diagnostic coding for billing and insurance
purposes is often handled by professional medical coders who must
explore a patient's extensive clinical record before assigning the
proper codes. So while electronic health records (EHRs) should be
adopted by most medical institutions within the next several years,
largely due to the provisions of HITECH under the American Recovery
and Reinvestment Act \citep{Blumenthal2009}, there has been little
movement forward in automating medical coding.

@@

While there have been some attempts to automatically classify groups
of patients as a potential preliminary step to ICD-9 code assignment
\citep{Ruch2008,FreitasJunior2006,RibeiroNeto2001,Brown2006}, fully
automatic assignment of ICD-9 codes to medical text became a more
prevalent research topic only in the last few years. A subset of earlier
work proposed various methods on small corpora, based on a few specific
diseases \citep{Rao2003} but the most recent and promising work on
the subject was inspired by the 2007 Medical NLP Challenge: \textquotedblleft{}International
Challenge: Classifying Clinical Free Text Using Natural Language Processing\textquotedblright{}
(website). Most of the classification strategies included word matching
and rule-based algorithms. \citep{Goldstein2007,Crammer2007,Farkas2008}.
The data set given to the participants consisted only of documents
that were 1-2 lines each and all of the documents were radiology reports
- clearly limiting the scope of potential ICD-9 codes which could
be assigned. The only paper which has attempted to work with a document
scope as large as ours was the 2008 Lita et al publication \citep{Lita2008}.
Lita proposed support vector machine and bayesian ridge regression
methods to assign appropriate labels to the documents but did not
utilize the ICD-9 hierarchy to leverage more comprehensive predictions.

@@

\section{Background}

\subsection{Data}

Our data set was gathered from the clinical data warehouse of NewYork
- Presbyterian Hospital. The data consisted of free-text discharge
summaries and their respective ICD-9 codes. A discharge summary is
a clinical report prepared by a physician or other health professional
at the conclusion of a hospital stay or series of treatments. The
note outlines the patient\textquoteright{}s chief complaint, diagnostic
findings, therapy administered, patient\textquoteright{}s response
to the chosen therapy, the treatment plan and the recommendations
upon discharge. The ICD-9 codes used to structure the discharge summary
data are part of a controlled terminology which is the international
standard diagnostic classification for epidemiological, health management,
and clinical purposes (http://www.who.int/classifications/icd/en/).
The codes are classified in a rooted-tree structure, with each edge
representing an is-a relationship between parent and child, such that
the parent diagnosis subsumes the child diagnosis. In the hospital,
ICD-9 codes are generated manually by trained medical coders, who
review all the information in the discharge summary. For the purposes
of sLDA, ICD-9 codes will be used as labels for discharge summaries.

\subsection{Pre-Processing}

Patient discharge summaries and their associated ICD-9 diagnoses are
stored in two different places in the NewYork - Presbyterian data
warehouse, and so had to be linked together before being fed into
the sLDA algorithm. Each discharge note and set of diagnoses were
assigned a patient unique identifier (PUID) and a visit unique identifier
(VUID), allowing the two types of data to be linked.

Preparation of the diagnostic codes involved inference over the ICD-9
hierarchy. The is-a relationships of the hierarchy allowed us to make
two important assumptions. First, if a diagnosis was observed to be
present, all of its ancestors could be assumed to be present as well
(e.g., if a patient had malignant hypertension, it could be assumed
that they also had essential hypertension. Second, if a diagnosis
was observed to be absent, it could be assumed that all of its descendants
were also absent (e.g. if a patient did not have essential hypertension,
it could be assumed that they did not have malignant hypertension).
Unfortunately, ICD-9 code observations never include observations
of disease absence. ICD-9 codes are only documented when the condition
is observed to be present. Additionally, ICD-9 codes are known to
have relatively low sensitivity; conditions that are present are often
not documented in a set of ICD-9 codes (Surjan, 1999). Given these
facts, we made the following assumptions regarding each visit: recorded
diagnoses and their ancestors were labeled as true; diagnoses that
were observed at some time for a patient but not at the current visit
were labeled as unobserved; and diagnoses that had never been listed
for a patient were labeled as false for all of that patient\textquoteright{}s
visits. This last assumption captures the belief that parts of the
ICD-9 hierarchy that are never observed for a particular patient are
almost certain to be false. Additionally, for computational purposes,
we decided not to include an ICD-9 code at all if neither it nor one
of its descendants had been assigned to a patient in any of the records
in our dataset.

\subsection{ICD-9 Code Hierarchy}

@@

%
\begin{figure}[h]
%tbp] %  figure placement: here, top, bottom, or page
 \centering \includegraphics[scale=0.3]{Graphical_Model2} \caption{adapted sLDA model}

Given the number of topics, K, and broad gamma priors on hyperparameters,
the generative process is as follows: 
\begin{enumerate}
\item For each topic, $k$:

\begin{enumerate}
\item Draw a distribution over words $\phi_{k}\sim Dir_{V}\left(\mathbf{1},\gamma\right)$ 
\end{enumerate}
\item For each ICD9 code, $c$, at all levels in the tree, $l$:

\begin{enumerate}
\item Draw regression coefficient $\eta_{i_{l,c}}\mid\sigma\sim\mathcal{N}_{K}\left(-\mathbf{1},\sigma\right)$ 
\end{enumerate}
\item Draw a prior over topic proportions $\beta\mid\alpha'\sim Dir_{K}\left(\mathbf{1},\alpha'\right)$ 
\item For each document, $d$:

\begin{enumerate}
\item Draw topic proportions $\theta_{d}\mid\beta,\alpha\sim Dir_{K}\left(\beta,\alpha\right)$ 
\item For each word, $n$:

\begin{enumerate}
\item Draw topic assignment $z_{n,d}\mid\theta_{d}\sim Mult_{K}\left(\theta_{d}\right)$ 
\item Draw word $w_{n,d}\mid z_{n,d},\beta_{1:K}\sim Mult_{V}\left(\beta_{z_{n}}\right)$ 
\end{enumerate}
\item For each level of the ICD-9 code tree, $l$:

\begin{enumerate}
\item For each ICD-9 code at this level, $c$:

\begin{enumerate}
\item Draw a latent variable \[
a_{d,i_{l,c}}\sim\begin{cases}
\mathcal{N}\left(\bar{z}^{T}\eta_{i_{l,c}},1\right), & y_{parent_{l,c}}=1\\
trunc\mathcal{N}^{-}\left(\bar{z}^{T}\eta_{i_{l,c}},1\right), & y_{parent_{l,c}}=-1\end{cases}\]
 where $\bar{z}=N^{-1}\sum_{n=1}^{N}z_{n}$ and $\mathcal{I}=\left\{ i_{0},i_{1},...,i_{\mathcal{L}}\right\} $
and $i_{l}=\left\{ i_{l,0},i_{l,1},...,i_{l,C_{l}}\right\} $
\item Draw a response variable $y_{d,i_{l,c}}\mid a_{d,i_{l,c}}\sim\begin{cases}
\begin{array}{c}
1,\\
-1,\end{array} & \begin{array}{c}
a_{d,i_{l,c}}>0\\
otherwise\end{array}\end{cases}$
\end{enumerate}
\end{enumerate}
\end{enumerate}
\end{enumerate}
The generative model for the ICD-9 codes is equivalent to a probit
regression model. In our case, the regression is conditional on the
parents according to the constraints of the ICD-9 code tree. The latent
variable utilized here is also known as an auxiliary variable.

Given an observation of a set of ICD-9 codes and a document, the posterior
distribution for the latent variables is given by Equation .

\begin{equation}
p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma\mid w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)\label{eq:Posterior}\end{equation}

\[
=\frac{p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma,w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)}{\int_{\theta,\phi,a,\eta,\alpha,\alpha',\beta,\gamma}\sum_{z}p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma,w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)}\]

The denominator for this distribution is the marginal probability
of the data and cannot be solved in closed form. This is often the
case in evaluating posterior distributions of non-trivial probabilistic
models. We will appeal to one of the common methods for approximating
posterior distributions in the face of intractable normalization factors:
Markov chain Monte Carlo (MCMC) sampling. Since in this model it is
possible to sample from the conditional distributions for all variables
we will use the Gibbs sampling algorithm to obtain an approximation
to this posterior. In particular, we will derive a collapsed Gibbs
sampler.

We derive a collapsed Gibbs sampler for this model, marginalizing
$\theta_{1:D}$ and $\phi_{1:K}$. For details regarding collapsing
in LDA models see \citet{Griffiths04}. To derive the Gibbs sampler
we evaluate the individual conditional probability distributions for
all unobserved variables.

\subsubsection{$p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)$}

For the purposes of sampling, we will be able to derive a representation
of the joint distribution isolating a particular latent variable,
$z$, for a word instance, $n$, in a document instance, $d$. The
conditional probability with respect to this latent variable is proportional
to the joint distribution of its markov blanket up to a constant.

\begin{equation}
p\left(z_{d,n}\mid\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\propto p\left(z_{d,n},\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\end{equation}

Due to the factorization of this model, we can rewrite the joint distribution
as the following:

\begin{equation}
\propto\prod_{i_{l,c}\in\mathcal{I}}p\left(a_{i_{l,c}}\mid\mathbf{z},\eta_{i_{l,c}}\right)p\left(z_{d,n},\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\alpha,\beta,\gamma\right)\end{equation}

\begin{equation}
\propto\prod_{i_{l,c}\in\mathcal{I}}\exp\left\{ -\frac{\left(\bar{z}^{T}\eta_{i_{l,c}}-a_{i_{l,c}}\right)^{2}}{2}\right\} p\left(a_{i_{l,c}}\mid\eta_{i_{l,c}}\right)\left(n_{d,\left(\cdot\right)}^{k,-\left(d,n\right)}+\alpha\beta_{k}\right)\frac{n_{\left(\cdot\right),w_{d,n}}^{k,-\left(d,n\right)}+\gamma}{\sum_{v=1}^{V}\left(n_{\left(\cdot\right),w_{d,n}}^{k,-\left(d,n\right)}+\gamma\right)}\label{eq:z-likelihood}\end{equation}

Here, $n_{d,v}^{k,-\left(d,n\right)}$ represents the count of word
$v$ in document $d$ assigned to topic $k$ omitting the $(d,n)^{th}$
word count.

Given Equation \ref{eq:z-likelihood}, $p\left(z_{d,n}\mid\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)$
can be sampled through enumeration.

\subsubsection{$p\left(\eta_{i_{l,c}}\mid\mathbf{z}_{1:D},\mathbf{a};\sigma\right)$}

Given that $\eta_{i_{l,c}}$ and $a_{d,i_{l,c}}$ are distributed
normally, this posterior distribution is also normal. We evaluated
the model over various values of $\sigma$ where $\sigma=\left\{ 0.01,0.1,0.25,1,2\right\} $.

\begin{equation}
p\left(\eta_{i_{l,c}}\mid\mathbf{z}_{1:D},\mathbf{a};\sigma\right)=\mathcal{N}\left(\eta_{i_{l,c}}\mid\hat{\mu}_{i},\hat{\Sigma}_{i}\right)\end{equation}

\[
\hat{\mu}_{i}=\hat{\mathbf{\Sigma}}_{i}\left(-\mathbf{1}\sigma^{-1}+\bar{\mathbf{Z}}^{T}\mathbf{a}_{\left(\cdot\right),i_{l,c}}\right)\]

\[
\hat{\Sigma}_{i}^{-1}=\mathbf{I}\sigma^{-1}+\bar{\mathbf{Z}}^{T}\bar{\mathbf{Z}}\]

\subsubsection{$p\left(a_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)$
and \textmd{$p\left(y_{m,i}\mid\mathbf{a}\right)$}}

In the augmented probit regression model, the posterior distribution
of $a_{i_{l,c}}$ is distributed according to a truncated normal distribution
where the response variable is observed.

\begin{equation}
p\left(a_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)=\begin{cases}
trunc\mathcal{N}^{+}\left(a_{d,i_{l,c}}\mid\eta_{i}^{T}\bar{z,}\mathbf{1},y_{d,i_{l,c}}\right) & if\quad y_{d,i_{l,c}}=1\\
trunc\mathcal{N}^{-}\left(a_{d,i_{l,c}}\mid\eta_{i}^{T}\bar{z,}\mathbf{1},y_{d,i_{l,c}}\right) & if\quad y_{d,i_{l,c}}=-1\end{cases}\end{equation}

However, if $y_{d,i_{l,c}}$ is unobserved then $a_{d,i_{l,c}}$must
be sampled jointly with $y_{d,i_{l,c}}$ to ensure that the Markov
chain is ergodic. Suppose that $a_{d,i_{l,c}}$ is sampled to have
a negative value and $y_{d,i_{l,c}}$ is apporopriately sampled at
-1. Although there exist valid states where $a_{d,i_{l,c}}>0$ and
$y_{d,i_{l,c}}=1$, they will never be reached by such a Markov chain
since $p\left(a_{d,i_{l,c}}<0\mid y_{d,i_{l,c}}=-1\right)=1$ and
$p\left(y_{d,i_{l,c}}=-1\mid a_{d,i_{l,c}}<0\right)=1$. Therefore,
to ensure ergodicity, $a_{d,i_{l,c}}$and $y_{d,i_{l,c}}$ must be
sampled from the joint distribution as shown in Equation \ref{eq:Probit-Joint}.

\begin{equation}
p\left(a_{d,i_{l,c}},y_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y}_{-\left(d,i_{l,c}\right)},\mathbf{\eta}\right)\propto p\left(y_{i_{l,c}}\mid\mathbf{a},\mathbf{y}_{-\left(l,c\right)}\right)p\left(a_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)\end{equation}

\begin{equation}
p\left(y_{i_{l,c}}\mid\mathbf{a},\mathbf{y}_{-\left(l,c\right)}\right)=\delta\left(sign\left(a_{d,i_{l,c}}\right)=y_{i_{l,c}}\right)p\left(y_{i_{l,c}}\mid y_{parents_{l,c}}\right)\prod_{i_{\hat{l},\hat{c}}\in children_{l,c}}p\left(y_{i_{\hat{l},\hat{c}}}\mid y_{i_{l,c}}\right)\end{equation}

\begin{equation}
p\left(y_{i_{l,c}}=-1\mid y_{parent{}_{l,c}}\right)=\begin{cases}
1, & y_{parent_{l,c}}=-1\\
0.5, & y_{parent_{l,c}}=1\end{cases}\end{equation}

\[
p\left(a_{d,i_{l,c}},y_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y}_{-\left(d,i_{l,c}\right)},\mathbf{\eta}\right)\]

\begin{equation}
=\begin{cases}
\mathcal{N}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right)p\left(y_{d,i_{l,c}}\mid a_{d,i_{l,c}}\right), & y_{parent_{l,c}}=1,\forall y_{i_{\hat{l},\hat{c}}}\in y_{children_{l,c}},y_{i_{\hat{l},\hat{c}}}=-1\\
trunc\mathcal{N}^{-}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right),\delta\left(y_{d,i_{l,c}}=-1\right), & y_{parent_{l,c}}=-1\\
trunc\mathcal{N}^{+}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right)\delta\left(y_{d,i_{l,c}}=1\right), & \exists y_{i_{\hat{l},\hat{c}}}\in y_{children_{l,c}}\setminus y_{i_{\hat{l},\hat{c}}}=1\\
0 & otherwise\end{cases}\end{equation}

where $\mathbf{Y}_{-\left(d,i_{l,c}\right)}$ denotes all of the response
variables excluding the response variable being sampled.

\subsubsection{$p\left(\beta\mid\mathbf{z},\alpha',\alpha\right)$}

In our model, we place a hierarchical Dirichlet prior over topic assignments.
This prior shares many features with the hierarchical Dirichlet process
and inference over this distribution proceeds in a very similar fashion.
This flexible distribution allows for an asymmetric prior over document
level distributions over topics \citet{WallachMiMc2009}.

Posterior inference was performed using the {}``direct assignment''
method of \citet{TehJorBea2006}.

\begin{equation}
\beta\sim Dir\left(m_{\left(\cdot\right),1},m_{\left(\cdot\right),2},\ldots,m_{\left(\cdot\right),K}\right)\end{equation}

\begin{equation}
p\left(m_{d,k}=m\mid\mathbf{z},\mathbf{m}_{-\left(d,k\right)},\beta\right)=\frac{\Gamma\left(\alpha\beta_{k}\right)}{\Gamma\left(\alpha\beta_{k}+n_{d,k}\right)}s\left(n_{d,k},m\right)\left(\alpha\beta_{k}\right)^{m}\end{equation}

where $s\left(n,m\right)$ represents stirling numbers of the first
kind.

\subsubsection{$p\left(\alpha;\lambda\right)$, $p\left(\alpha';\lambda\right)$,
$p\left(\beta;\lambda\right)$}

All hyperparameters were given broad gamma priors $\left(\lambda=\left\{ shape=2,scale=1000\right\} \right)$
and sampled via the Metropolis-Hastings algorithm.

\subsection{Prediction}

\subsection{Related Work}

Latent dirichlet allocation (LDA) is a generative probabilistic model
of corpora that represents documents as a mixed membership bag-of-words.
Also known as topic models, these models infer the latent structure,
or topics, of documents in a corpus. Each document is represented
as a collection of words, generated from a set of topic assignments
(one for each word), where each topic assignment is drawn from a distribution
over topics \citep{Blei2003}.

@@

Other models - predicting document links, other supervised latent
variable models
